---
title: "1-Define Data"
author: "Verooo"
date: "2024-11-01"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
klippy::klippy(color = 'darkred')
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

[**Defining Data**]{.underline}

# Before the start

## Clean workspace

```{r}
# start with clean workspace
rm(list = ls())

```

## Packages

```{r}
# install.packages('tidyverse')
require(tidyverse)

# install an old version of RSiena including sienaRI
# remove.packages("RSiena") #use this to remove the current version of RSiena
#devtools::install_version("RSiena", version = "1.3.14", repos = "http://cran.us.r-project.org") # This installs an older version

# Run R as administrator when you install the older version!!
  
# You should end up with version 1.3.14 loaded in instead of 1.4.7, 
# but then make sure there isn't a install.packages("RSiena") function running somewhere, because then it will reinstall the newest version

```

## Functions

### Question

**HOW do I get the discipline of political science in there too?**

**How can I find out what the name for the discipline of political science in the original data frame (e.g. whether it is called polsci, political science, or politicalscience)?**

**I don't know how to view the data set BEFORE we include only those scholars that have discipline = sociology. I cannot view the data**

I tried removing the discipline = "sociology" part in the fcolnet and all following functions that select data BUT SINCE 'discipline' is also relevant in the rest of the functions' code, it does not end up doing what I hoped (having a dataset with scholars of no specific discipline because then I could hopefully check how the department is named in the data and then find out how to add that to the data set initially).

```{r}

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}


# before my edits:
# f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
#     wave = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023)))

#after taking out discipline= "sociology" in this first part    
f_pubnets <- function(df_scholars = df, list_publications = publications, affiliation = "RU",
    wave = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(wave)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(wave)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% wave[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(wave)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(wave)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(wave)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}
```

```{r}
view(f_pubnets)
```

## more functions

```{r}

#prior to my edits:
# fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", wave = list(c(2019, 2023)), type = c("first")) 

#after taking out discipline= "sociology" in the first part 
fcolnet <- function(data = scholars, university = "RU", wave = list(c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwave <- length(wave)
    nets <- array(0, dim = c(nwave, length(ids), length(ids)), dimnames = list(wave = 1:nwave, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwave) {
            df_works_w <- df_works[df_works$works_year >= wave[[j]][1] & df_works$works_year <= wave[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwave) {
            df_works_w <- df_works[df_works$works_year >= wave[[j]][1] & df_works$works_year <= wave[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwave) {
            df_works_w <- df_works[df_works$works_year >= wave[[j]][1] & df_works$works_year <= wave[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
    
}

```

## load data

```{r}
scholars <- fload("./data/processed/scholars_20240924.rda")

# save the output of your function

# before my edits:
# test <- fcolnet(data = scholars, university = "RU", discipline = "sociology", wave = list(c(2019, 2023)), type = c("first"))

# after taking discipline= "sociology" out
# test <- fcolnet(data = scholars, university = "RU", wave = list(c(2019, 2023)), type = c("first"))

# here is where the error occurs due to taking out the discipline = "sociology" in multiple prior parts of the code 
```

------------------------------------------------------------------------

# Define Data

```{r}
wave <- test$nets[1, , ]

# let us put the diagonal to zero
diag(wave) <- 0
# we want a binary tie (not a weighted tie)
wave[wave > 1] <- 1

# put the nets in an array
net_soc_array <- array(data = c(wave), dim = c(dim(wave), 2)) 
str(net_soc_array)  #now we basically restructured this 
# output: num [1:50, 1:50, 1:2] 0 0 0 0 0 0 0 0 0 0

# since I only take one wave should I replace the 2 with a 1 here? 
net_soc_array1 <- array(data = c(wave), dim = c(dim(wave), 1)) 
str(net_soc_array1)  #now we basically restructured this 
# output: num [1:50, 1:50, 1] 0 0 0 0 0 0 0 0 0 0
```

### **QUESTION:**

**What's the difference between putting the 2 and the 1 at the dimensions. I feel like one is appropriate since I am only looking at one wave, not two, but since the output is among the 'values' not 'data', I cannot view the variable net_soc_array(1) nor "visualize" what that means in terms of its structure. I don't really know what the difference would be between the 1:2 and the 1 in the outputs, structurally.**

## Examine structure

```{r}
dim(test$nets)
```

not using net_soc_array for now since I am unsure about its use due to question above

```{r}
# dependent
    # not using net_soc_array for now since I am unsure about its use due to question above
# net <- sienaDependent(net_soc_array)


# dependent 
 nets <- test$nets

 
df_ego_demo <- test$data$demographics
df_ego_oa <- test$data$scholars_oa
df_ego_works <- test$data$works


# DEMO
  str(df_ego_demo, 1) 
  
  df_ego_demo <- do.call(rbind.data.frame, df_ego_demo)
  df_ego_demo[[1]]

# OA
  #str(df_ego_oa, 1)
  df_ego_oa[[1]]
  df_ego_oa <- do.call(rbind.data.frame, df_ego_oa)


# --> EGO DATAFRAME with demo and oa 
  ?join.data.frame
  df_ego <- left_join(df_ego_demo, df_ego_oa, by = join_by(au_id == id))
  
  rownames(nets[1,,]) == df_ego$au_id # checking if the author id's are the same as in df_ego in nets dataframe 
  # --> yes, the correct ones!!!


df_ego_works[[1]]$author 
#authors of all the works that first scholar (here: Ana Macanovic) is part of 


```

## Gender

```{r}
df_ego$Naam # list of all names 

df_ego$gender <- c(0,1,0,1,0,0,1,1,0,0,
                   0,1,1,0,1,0,1,1,1,1,
                   0,1,1,0,1,1,0,1,0,1,
                   0,1,0,0,0,0,0,0,1,0,
                   0,1,1,1,0,1,0,1,0,0) 
# manually filled in gender 0 = woman, 1 = man, based on majority given in NL voornamen bank 

# check if it went smoothly 
view(df_ego[c('Naam', 'gender', 'au_id')])

#covariate 
# gender <- coCovar(df_ego$gender)
```

## H-index

Inspired by Lucan's code:

```{r, include=FALSE}
# Add h-index and i10-index columns to df_ego
df_ego <- df_ego %>%
  mutate(h_index = NA_real_, i10_index = NA_real_)

# Get H-index and i10-index 
get_openalex_metrics <- function(openalex_id) {
  url <- paste0("https://api.openalex.org/authors/", openalex_id)
  author_data <- tryCatch({
    jsonlite::fromJSON(url)
  }, error = function(e) {
    message(paste("Error fetching data for OpenAlex ID:", openalex_id))
    return(NULL)
  })
  
  if (!is.null(author_data)) {
    h_index <- author_data$summary_stats$h_index
    i10_index <- author_data$summary_stats$i10_index
    works_count <- author_data$works_count
    cited_by_count <- author_data$cited_by_count
    return(list(h_index = h_index, i10_index = i10_index, works_count = works_count, cited_by_count = cited_by_count))
  } else {
    return(list(h_index = NA, i10_index = NA, works_count = NA, cited_by_count = NA))
  }
}

# Loop for each scholar 
for (i in 1:nrow(df_ego)) {
  openalex_id <- df_ego$au_id[i]  # assuming au_id is the OpenAlex ID in df_ego
  metrics <- get_openalex_metrics(openalex_id)
  df_ego$h_index[i] <- metrics$h_index
  df_ego$i10_index[i] <- metrics$i10_index
  df_ego$works_count[i] <- metrics$works_count
  df_ego$cited_by_count[i] <- metrics$cited_by_count
}

# View updated df_ego with h-index and i10-index
view(df_ego[c('au_id','Naam','gender','h_index','i10_index')])


#covariate 
# h_index <- coCovar(df_ego$h_index)
```

```{r}

view(data('discipline'))
```
