---
title: "draft-data"
author: "Verooo"
date: "2024-11-01"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Data Draft

Information about data set:

-   Data set provided by Jochem

-   collaborations of scholars in NL

-   time frame: ?

    -   timeframe of focus

## DV: Collaboration between scholars 

## Covariate 1: Gender

-   assumed binary for simplicity and ethical considerations of assuming gender based on other characterstics

-   Names of scholars manually checked in Dutch Voornamenbank website **(NAME WEBSITE)**

-   with 0 = female and 1 = male.

-   

## Covariate 2: Prestige

-   indicated by H-Index: and i10 index

-   definitions indeces:

-   

-   number of citations and publications webscraped through OpenAlex **[source]**

-   h index and i10 index calculated based on these counts

```{=html}
<!-- -->
```
-   
