---
title: "Draft-Theory"
author: "Verooo"
date: "2024-10-31"
output: html_document
bibliography: references.bib
csl: apa.csl
---

```{=html}
<style>
body {
text-align: left;
line-height: 1.5;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  line-height: 1;
}

</style>
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Theory Draft 2 

## **1.0.1 Theory and Hypotheses**

Understanding collaboration patterns in academia requires examining how individual preferences, structural factors, and institutional norms interact to shape scholarly networks. Two key mechanisms guide this inquiry: **homophily**, the tendency to form ties with similar others (@McPherson2001), and **preferential attachment**, where individuals gravitate toward high-prestige collaborators (@Merton1968). Additionally, the concept of **upward mobility** highlights how strategic collaborations with prestigious scholars can counterbalance preferences based on similarity, such as gender homophily.

The study tests four hypotheses across the Political Science and Sociology departments at Radboud University to explore how these mechanisms interact to influence collaboration patterns.

### **H1: Gender Homophily in Collaboration Networks**

Homophily explains the widespread tendency of individuals to associate with similar others, a phenomenon deeply embedded in academic networks (@McPherson2001). In this context, **gender homophily** manifests as a preference for same-gender collaborations, resulting in gendered clustering within scholarly networks. Research highlights that men frequently collaborate with other men, and women with other women, perpetuating disparities in visibility, resource access, and academic advancement (@Horta2022; @Kwiek2021).

This hypothesis builds on evidence of gendered collaboration patterns in academia and investigates whether these patterns persist across diverse academic contexts.

***H1:** Across the collaboration networks of both departments, scholars are more likely to collaborate with same-gender peers than with different-gender peers, indicating the presence of gender homophily.*

### **H2: Variation in Gender Homophily Between Departments**

The extent of gender homophily may vary by discipline due to differences in gender representation and cultural norms. Sociology, often seen as more gender-balanced or even female-dominated, may facilitate cross-gender collaborations more than Political Science, which remains a male-dominated field associated with traditional masculine traits such as assertiveness and rationality (@Kwiek2021). These cultural perceptions influence collaboration preferences, potentially reducing gender homophily in disciplines where gender parity is normalized.

This hypothesis tests whether departmental norms and gender dynamics shape the degree of homophily observed in collaboration networks.

***H2:** Scholars in the Sociology Department exhibit weaker gender homophily in their collaborations than those in the Political Science Department.*

### **H3: Structural Network Effects and Gender Homophily**

Collaboration networks are also shaped by structural effects such as **reciprocity** (mutual collaboration), **indegree popularity** (preference for well-connected individuals), and **outdegree activity** (individuals actively seeking collaborators). These mechanisms reinforce broader social dynamics in scholarly networks, often amplifying existing tendencies such as gender homophily (@Snijders2010). For example, well-connected scholars may still attract same-gender collaborators due to persistent societal norms or unconscious biases, even as their popularity grows.

This hypothesis examines whether structural effects enhance collaboration patterns while leaving underlying gender preferences intact.

***H3:** Structural network effects, such as reciprocity and indegree popularity, amplify collaboration tendencies but do not diminish the effect of gender homophily.*

### **H4: The Influence of Prestige on Gender Homophily**

While homophily shapes collaboration patterns, prestige can alter these dynamics by creating opportunities for **upward mobility**. Prestige, often measured by the Hirsch-Index, serves as a powerful signal of influence and impact in academia (@Bornmann2020).Scholars seeking career advancement may prioritize collaborations with prestigious colleagues, even if it means collaborating across gender lines(@Broido2019). However, as women and other underrepresented groups gain visibility and prestige, their collaboration opportunities expand, potentially mitigating homophily effects (@Knobloch2013). Hence, the last hypothesis explores whether the pursuit of prestige mitigates the influence of gender homophily in academic collaboration networks.

***H4:** Gender homophily is less influential in collaboration decisions when the potential collaborator has a higher Hirsch-Index.*

------------------------------------------------------------------------

# Theory Draft 1

### Theory

Understanding collaboration patterns in academia id crucial for unpacking the mechanisms behind the formation of scholarly networks. One key principle that shapes these networks is **homophily**, the tendency of individuals to associate with others who are similar to them (@McPherson2001). In the context of academic collaborations, **gender homophily**—the preference for scholars to collaborate with others of the same gender—has been widely observed. Research consistently shows that men tend to collaborate more frequently with other men, and women with other women ( @Horta2022; @Kwiek2021). However, the degree and dynamics of this homophily can vary across disciplines and contexts, making it an interesting field for investigation.

#### Hypothesis 1: Gender Homophily in Collaboration Networks

The first hypothesis centers on the presence of gender homophily across two academic departments: Political Science and Sociology at Radboud University. Based on the homophily principle, I expect scholars to prefer same-gender collaborators, leading to gendered clustering in co-authorship networks.

-   ***H1**: Across the collaboration networks of both departments, scholars are more likely to collaborate with same-gender peers than with different-gender peers, indicating the presence of gender homophily.*

This hypothesis aligns with prior studies that highlight gender homophily as a key feature of collaboration networks in academia (@Kwiek2021). However, the extent to which this tendency manifests may depend on the structural characteristics of each department, as well as broader societal norms and stereotypes.

#### Hypothesis 2: Variation in Gender Homophily Between Departments

While gender homophily is a general phenomenon, its intensity may vary across disciplines due to their differing gender dynamics and societal perceptions. Sociology, often perceived as a more gender-balanced or even female-leaning field, may exhibit weaker gender homophily compared to Political Science, which is often associated with traditionally masculine traits such as rationality and assertiveness. These stereotypes may influence collaboration patterns, with scholars in more male-dominated fields showing stronger gender homophily (@Kwiek2021).

-   ***H2**: Scholars in the Sociology Department exhibit weaker gender homophily in their collaborations than those in the Political Science Department.*

This hypothesis builds on the idea that the gender composition and cultural norms of academic disciplines shape collaboration preferences. In sociology, where gender parity is more pronounced, collaboration across genders may be more normalized, reducing homophilic tendencies.

#### Hypothesis 3: The Influence of Prestige on Gender Homophily

While homophily can drive collaboration patterns, prestige also plays a critical role. Scholars are often drawn to collaborate with those who have high academic prestige, typically measured through metrics such as publication in high-impact journals or h-index scores (@Bornmann2020). This "prestige effect" may weaken the influence of gender homophily, as the opportunity to collaborate with a prestigious scholar may outweigh same-gender preferences.

-   ***H3**: Gender homophily is less influential in collaboration decisions when the potential collaborator has a higher Hirsch-Index.*

This hypothesis posits that prestige (higher h-index) serves as a countervailing force to homophily. Collaboration with high-prestige individuals is a strategic choice, driven by the potential for increased visibility and career advancement, which may diminish the salience of gender in collaboration decisions.

### Theoretical Framework

The theoretical framework draws on both **homophily theory** and the concept of **preferential attachment** in academic networks. The homophily principle explains the tendency for scholars to form ties with others who share their gender, while preferential attachment highlights the appeal of prestigious collaborators ( @Merton1968; @Broido2019). Together, these mechanisms shape thr structure and evolution of collaboration networks.

Prior research suggests that men, particularly in male-dominated fields, are more likely to form same-gender ties, reinforcing gender inequalities in academia (@Kwiek2021). However, as women gain visibility and prestige, their collaboration opportunities expand, potentially mitigating homophily effects (@Knobloch2013). Understanding these dynamics requires a nuanced approach that accounts for both individual preferences and structural network constraints.

By addressing these hypotheses, this study contributes to the broader literature on academic collaboration and gender dynamics, offering insights into how societal norms and institutional contexts influence the formation of scholarly networks.

# prior notes:

## gender homophily in scholars' collaboration

-   often: Males publishing with males, females publishing with females (@Horta2022)

-   **Hypothesis 1**: across the collaboration networks of both departments (pol. sci & soc.) scholars prefer to collaborate with same-sex scholars over different-sex scholars (gender homophily is present).

    -   [RQ1: To what extent does gender homophily occur across all co-publications of the staff of the Sociology Department and Political Science Department at Radboud University? descriptive, at network level]\

[\*\*look at descriptives: how many men how many women in each department]\*\*

-   Due to bias that social jobs are more feminine and politics are related to rationality, which is often stereotyped to be masculine: Expect less gender homophily in sociology than in pol.sci. department.

-   (logical error: yeah but that only makes sense if men are the only ones initiating and breaking ties, right?)

<!-- -->

-   **Hypothesis 2**: between the collaboration networks of both departments (pol. sci & soc.) scholars in the sociology department prefer to collaborate with same-sex scholars less strongly than scholars from the pol. sci. department

    -   [RQ2: To what extent do the two departments differ in their gender homophily?\*descriptive, at the dyad-level (after controlling for structural network effects and covariate-effects].\*

<!-- -->

-   *hypothesis for your RSiena models: whether within a dyad men tend to (or 'prefer') work with men and women with women, (controlled for structural network effects).*

    -   **confusion: where would the comparison between depatments be here? this wold fit rq one more, but that one is supposed to be on the network level but this is about dyads...**

<!-- -->

## h-index and collaboration

-   **Hypothesis 3:** On the ego level, the preference of *men to work with men and women with women, (controlled for structural network effects), is less influential in the choice to collaborate if the alter is someone who published in a high-prestige journal.*

    -   [RQ3: How does gender homophily in collaborations between staff differ depending on scholars having published in low vs. high prestige journals? ego-level] *idea: whether prestige (of ego/alter/dyad) impacts gender homophily in some way.*

# Written references (remove in final paper)

Bornmann, Lutz, Christian Ganser, and Alexander Tekles. 2021. “Simulation of the h Index Use at University Departments Within the Bibliometrics-Based Heuristics Framework: Can the Indicator Be Used to Compare Individual Researchers?” Journal of Informetrics 16 (1): 101237. <https://doi.org/10.1016/j.joi.2021.101237>.

@Bornmann2021

Bornmann, Lutz, Christian Ganser, Alexander Tekles, and Loet Leydesdorff. 2020. “Does the h αα-Index Reinforce the Matthew Effect in Science? The Introduction of Agent-Based Simulations into Scientometrics.” Quantitative Science Studies 1 (1): 331–46.

@Bornmann2020

Broido, A. D., & Clauset, A. (2019). Scale-free networks are rare. *Nature communications*, *10*(1), 1017. <https://doi.org/10.1038/s41467-019-08746-5>

@Broido2019

Merton, Robert K. 1968. “The Matthew Effect in Science: The Reward and Communication Systems of Science Are Considered.” *Science* 159 (3810): 56–63.

@Merton1968

Knobloch-Westerwick, Silvia, Carroll J Glynn, and Michael Huge. 2013. “The Matilda Effect in Science Communication: An Experiment on Gender Bias in Publication Quality Perceptions and Collaboration Interest.” *Science Communication* 35 (5): 603–25.

@Knobloch2013

Snijders, T. A., Van De Bunt, G. G., & Steglich, C. E. (2009). Introduction to stochastic actor-based models for network dynamics. *Social Networks*, *32*(1), 44–60. <https://doi.org/10.1016/j.socnet.2009.02.004>

@Snijders2010

------------------------------------------------------------------------

# References
