---
title: "Theory"
author: "Verooo"
date: "2024-11-23"
output: html_document
bibliography: references.bib
csl: apa.csl
---

```{=html}
<style>
body {
text-align: left;
line-height: 1.5;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  line-height: 1;
}

</style>
```
```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Theory & Hypotheses

Understanding collaboration patterns in academia requires examining how individual preferences, structural factors, and institutional norms interact to shape scholarly networks. Two key mechanisms guide this inquiry: **homophily**, the tendency to form ties with similar others (@McPherson2001), and **preferential attachment**, where individuals gravitate toward high-prestige collaborators (@Merton1968). Additionally, the concept of **upward mobility** highlights how strategic collaborations with prestigious scholars can counterbalance preferences based on similarity, such as gender homophily. The study tests four hypotheses across the Political Science and Sociology departments at Radboud University to explore how these mechanisms may interact to influence scholars' collaboration patterns.

## **H1: Gender Homophily in Collaboration Networks**

Homophily explains the widespread tendency of individuals to associate with similar others, a phenomenon deeply embedded in academic networks (@McPherson2001). In this context, **gender homophily** manifests as a preference for same-gender collaborations, resulting in gendered clustering within scholars' networks. Research highlights that men frequently collaborate with other men, and women with other women, perpetuating disparities in visibility, resource access, and academic advancement (@Horta2022; @Kwiek2021). Thus, the first hypothesis builds on evidence of gendered collaboration patterns in academia and investigates whether these patterns persist across diverse academic contexts.

-   ***H1:** Across the collaboration networks of both, the sociology and the political sciene department, scholars are more likely to collaborate with same-gender peers than with different-gender peers, indicating the presence of gender homophily.*

## **H2: Variation in Gender Homophily Between Departments**

The extent of gender homophily may vary by discipline duee to differences in gender representation and cultural norms. Sociology, often seen as more gender-balanced or even female-dominated, may facilitate cross-gender collaborations more than Political Science, which remains a male-dominated field associated with traditional masculine traits such as assertiveness and rationality (@Kwiek2021). These cultural perceptions influence collaboration preferences, potentially reducing gender homophily in disciplines where gender parity id normalized. The second hypothesis tests whether departmental norms and gender dynamics affect scholars' tendency for gender homophily in their collaborations.

-   ***H2:** Scholars in the sociology department exhibit weaker preferences for same-gender collaborators than those in the political science department.*

## **H3: Structural Network Effects and Gender Homophily**

Collaboration networks are also shaped by structural effects such as **reciprocity** (mutual collaboration), **indegree popularity** (preference for well-connected scholars), and **outdegree activity** (scholars actively seeking collaborators). These mechanisms reinforce broader social dynamics in scholarly networks, often amplifying existing tendencies such as gender homophily (@Snijders2010). For example, well-connected scholars may still attract same-gender collaborators due to persistent societal norms or unconscious biases, even as their popularity grows. Hence, the next hypothesis examines whether structural effects enhance collaboration patterns in scholars tie creations.

-   ***H3:** Gender homophily in collaboration between scholars persists after controlling for network structure (e.g., reciprocity, degree distribution).*

## **H4: The Influence of Prestige on Gender Homophily**

While homophily shapes collaboration patterns, prestige can alter these dynamics by creating opportunities for **upward mobility**. Prestige, often measured by the Hirsch-Index (h-index), serves as a powerful signal of influence and impact in academia (@Bornmann2020).Scholars seeking career advancement may prioritize collaborations with prestigious colleagues, even if it means collaborating across gender lines(@Broido2019). However, as women and other nderrepresented groups gain visibility and prestige, their collaboration opportunities expand, potentially mitigating homophily effects (@Knobloch2013). Hence, the last hypothesis explores whether high prestige mitigates the influence of gender homophily in academic collaboration networks.

-   ***H4:** Gender homophily is less influential in collaboration decisions when the potential collaborator has a higher Hirsch-Index.*

------------------------------------------------------------------------

# References
