---
title: "Discussion"
author: "Verooo"
date: "2024-11-23"
output: html_document
bibliography: references.bib
csl: apa.csl
---

```{=html}
<style>
body {
text-align: left;
line-height: 1.5;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  line-height: 1;
}

</style>
```
```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Discussion

This study investigates how co-authorship networks among scholars are influenced by gender, departmental affiliation, and academic prestige. By examining networks within the Political Science and Sociology departments at Radboud University, this study utilizes 'RSiena', simulating changes in social networks, to understand how these factors jointly shape scholars' collaboration patterns (@RSiena_manual, @RSiena_package. Scholars often prefer collaborators of the same gender, a phenomenon rooted in societal norms and perceived status differences ( @Horta2022 ). This is expected to result in gender clustering within academic networks, with men and women showing distinct preferences for same-gender collaborations (*Hypothesis 1).* Disciplines differ in their gender dynamics. Sociology, characterized by a more balanced gender representation, may facilitate cross-gender collaborations more so than the field of Political Science (@Kwiek2021). Hence, gender homophily was expected to be less influential in collaboration preferences of scholars in the sociology department than those in political science *(Hypothesis 2).\
*Furthermore, structural features of collaboration networks, such as reciprocity, indegree popularity, and outdegree activity, shape how collaboration ties form (@Snijders2010). These structural network effects may reinforce gender homophily in collaboration preferences in the creation of ties betwen scholars *(Hypothesis 3)*. Prestige, as measured by the Hirsch-Index, may mitigate gender homophily by increasing the appeal of high-status collaborators regardless of their gender (@Bellotti2022). This hypothesis examined whether strategic collaborations dilute gendered preferences in scholarly networks. Specifically it investigated whether gender homophily is less promnounced in scholars' preference to collaborate when potential collaborators have a higher H-index *(Hypothesis 4).*

## Situating Findings

### **Gender Homophily (H1)**

The results consistently support Hypothesis 1, showing significant evidence of gender homophily across all models. The parameter for same-gender collaboration (sameX) remains positive and significant, indicating that scholars are more likely to collaborate with same-gender peers than with different-gender peers. This aligns with previous research suggesting that gendered barriers in academia, whether structural or cultural, can lead to gender clustering in collaborative networks (@Horta2022; @Kwiek2021; @McPherson2001).\
Interestingly, the significance of gender clustering persists even after controlling for other covariates, such as department affiliation and h-index, demonstrating that it is a robust phenomenon in these networks. While this result is not unexpected, it emphasizes the enduring influence of gender dynamics on scholarly collaborations.

### **Departmental Differences in Gender Homophily (H2)**

Hypothesi 2 proposed that scholars in the sociology department exhibit weaker gender homophily than those in the political science department. While this hypothesis was not explicitly tested using an interaction term, comparisons across models suggest no substantial department-level differences in gender clustering. The same-gender collaboration parameter (sameX for gender) is nearly identical in models 2 and 3, indicating that the inclusion of department as a covariate does not significantly alter the pattern of same-gender collaborations.\
However, the department-related parameter for ego’s tendency to initiate collaborations (egoX for department) becomes significant in model 4, revealing a lower likelihood for political scientists to initiate collaborations compared to sociologists. This could imply department-level differences in collaboration norms or opportunities, which might indirectly influence gender homophily. An interaction term between department and gender would have been useful to further investigate these nuances. After all, hypothesis 2 was not supported. While scholars show a significant preference for collaborating within their own department, with sociologists more likely to initiate collaborations than political scientists, *no* substantial differences in gender homophily were observed in relation to scholars' departments.

### **Structural Network Effects and Gender Homophily (H3)**

In terms of structural network effects, reciprocity emerges as a consistently significant factor, indicating that collaborations are more likely to be mutual. This aligns with existing theories of dyadic reciprocity in networks, suggesting that once a collaboration tie is initiated, it is likely to be reciprocated (@Snijders2010). Interestingly, the outdegree parameter (baseline likelihood of initiating a tie) fluctuates in significance across models but remains consistently negative when significant, indicating that scholars generally show a low tendency to initiate collaborations. In line with Hypothesis 3, Gender homophily (sameX) remains significant even when controlling for these structural effects, reinforcing its importance. This finding implies that scholars' same-gender preferences in creating ties are not merely a byproduct of network structure but reflect deeper social or cultural dynamics.

### Prestige and Gender Homopgily (H4)

The results provide little to no support for Hypothesis 4, which proposed that gender homophily is less influential when potential collaborators have higher academic prestige (as measured by the h-index). The inclusion of h-index as a covariate in model 4 reveals that scholars with higher h-index scores are more likely to attract collaborations (significant positive alter effect, altX) but less likely to initiate them (significant negative ego effect, egoX). This is consistent with theories of preferential attachment and upward mobility, where prestigious scholars act as hubs, receiving collaboration requests rather than initiating them (@Broido2019; @Knobloch2013).\
However, since the the same-gender parameter (sameX) remained significant and largely unchanged, the influence of genderhomophily on scholars' decisions to collaborate is not substantially impacted by scholars' h-index. This finding suggests that while prestige plays a role in shaping collaboration networks, it does mitigate the tendency of scholars to prefer same-gender collaborations over different-gendered ones. Thus, while prestigious scholars are more sought after, collaboration preferences in scholars' networks remain influenced by gender-preferences.

## **Broader Implications**

The findings of this study highlight the interplay between structural, departmental, and individual factors in shaping co-authorship networks. Gender homophily appears to be a robust and enduring feature of these networks, resisting the influence of prestige and departmental context. This raises important questions about the structural barriers faced by scholars, particularly women, in accessing diverse collaboration opportunities.\
From a policy perspective, these results underscore the need for initiatives aimed at reducing gendered collaboration clustering. For example, mentorship programs or institutional incentives that encourage cross-gender and cross-department collaborations could help lessen these pattern. Additionally, the finding that prestigious scholars are less likely to initiate collaborations suggests that researchers early int heir career, particularly those from underrepresented groups, may face challenges in accessing these high-prestige hubs. Institutions should consider strategies to facilitate connections between early-career and high-prestige scholars to promote equity in collaboration opportunities.

## **Limitations and Future Research**

This study has several limitations that warrant consideration. First, the lack of interaction terms between gender and department or h-index limits the ability to make strong conclusions about how these factors jointly influence collaboration patterns. Future studies should include such interactions to better understand these dynamics. Second, while the h-index serves as a useful proxy for academic prestige, this study did not control for author characteristics such as age, which may account for some of the effects found. Besides, other measures of prestige or influence, such as grant funding or editorial roles, might yield additional insights. Finally, the examined sample is arguably small, and the generalizability of these findings is further limited to the specific departments and network structures examined here. \
Future research should explore longitudinal changes in collaboration patterns to assess whether gender homophily weakens over time, particularly as scholars gain prestige. Additionally, examining the role of institutional interventions, such as funding programs or diversity initiatives, could provide valuable insights into strategies for fostering more equitable collaboration networks.

## **Conclusion**

This study contributes to our understanding of the dynamics of co-authorship networks by highlighting the persistent influence of gender homophily in scholars' collaboration preferences, even in the presence of academic prestige and structural network effects. While prestigious scholars are more likely to attract collaborations, gendered patterns of collaboration endure. This underrscores the need for targeted interventions to address these disparities. By disentangling the effects of gender, department, and prestige, this research provides a foundation for future studies and practical efforts aimed at fostering more inclusive and diverse academic collaboration networks.

------------------------------------------------------------------------

# References
