---
title: "Model estimation"
author: "Verooo"
date: "2024-11-20"
output: html_document
bibliography: references.bib
csl: apa.csl
editor_options: 
  markdown: 
    wrap: 72
---

<html lang="en">

<head>

<meta charset="UTF-8">

<meta name="viewport" content="width=device-width, initial-scale=1.0">

<title>My Descriptives</title>

<link rel="stylesheet" href="styles.css">
<!-- Link to the CSS file I created for dark mode code chunks -->

</head>

</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Note to self. Keep in mind: \
**meaning of significance in RSiena** "It is important to mention that
significance in RSiena means something different than in conventional
statistics. In normal statistics, such as OLS-regression, significance
of estimates refers to the fact that they are not 0 (**RELEVANT?**). In
RSiena, however, it refers to the fact that the estimates were precisely
determined (Ripley et al. 2011). Going forward with this chapter,
significance will refer to this definition of significance." - Niels
Vullings

# Setup

Clean workspace. As always.

```{r clean}
rm(list = ls())
```

## Packages

```{r package, include=TRUE, hide=TRUE, eval=TRUE}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)


#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)


```

## Functions

Functions chunk:

```{r functions 1, hide=TRUE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}


# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions:

```{r functions 2, hide= TRUE}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

# Get Data

Load Scholar Data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
```

Save the output of the function.

```{r saving output}
# save the output of your function
test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
```

Load Ego Data

```{r df_ego data, include= FALSE}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

# Make Network Data

```{r RSiena network data}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

# put nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

```

------------------------------------------------------------------------

# Analysis

```{=html}
<!--
## TEST MODEL

### test 1: Define Siena Data & Variables
-->
```
```{r RSiena define, hide =TRUE, echo=FALSE, eval=FALSE}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena


# put h-index in object 
h_index <- df_ego$h_index
df_ego$h_index

# covariate 2
h_index_RS <- coCovar(h_index) # h_index as covariate in RSiena

# define data object 
mydata <- sienaDataCreate(nets, gender_RS, h_index_RS)

```

<!-- Print short descriptive report-->

```{r, hide =TRUE, echo=FALSE, eval=FALSE}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

#if there is not a folder called "results" in your repository: make one

```

```{r, hide =TRUE, echo=FALSE, eval=FALSE}
print01Report(mydata, modelname = "./results/try") # create file with report in results folder 
```

<!-- ### test 2: Make effects object/structure -->

```{r RSiena myeff, hide =TRUE, echo=FALSE, eval=FALSE}

# define effects structure 
myeff <- getEffects(mydata)

```

```{=html}
<!--

e.g. outdegree (density), called density, three things: evaluation,
endowment, and creation you can only include two of the three (tie
creation, breaking and maintaining)

The assumption is: mechanisms to make and break a a tie are the same so
WE ONLY USE THE EVALUATION part

All these effects are described in the manual of RSiena Chapter 12!

-->
```
```{=html}
<!--  
### Step X: Initial description
-->
```
```{=html}
<!--
Has a lot of information in there Also: how tie CHANGED Also: Jaccard
index (measure of stability) Also rate parameter and outdegree

<https://www.stats.ox.ac.uk/~snijders/siena/RSiena_Manual.pdf>

In the package RsienaTwoStep there are functions for the following
network statistics ***s*** :

degree: `ts_degree()`

reciprocity: `ts_recip()`

outdegree activity: `ts_outAct()`

indegree activity: `ts_inAct()`

outdegree popularity: `ts_outPop()`

indegree popularity: `ts_inPop()`

transitivity: `ts_transTrip()`

mediated transitivity: `ts_transMedTrip()`

transitive reciprocated triplets: `ts_transRecTrip()`

number of three-cycles: `ts_cycle3()`
-->
```
<!-- ### test 3: Add Effects -->

```{r effects, hide =TRUE, echo=FALSE, eval=FALSE}

# Null model: just myeff (just degree and reciprocity by default) # model0  #for this when you create Siena Data, do not add the gender variable 

myeff <- includeEffects(myeff, isolateNet, inPop, outAct) 
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collab with people who are popular
  # over time: inpop: ppl with lots of indegrees will havelot of indegrees

#in-degree related popularity effect (inPop) (earlier called popularity or popularity of alter effect), defined by the sum of the in-degrees of the others to whom i is tied,

# later (in population) outAct: ppl with lots of outdegrees will have lot
# of indegrees later


myeff <- includeEffects(myeff, egoX, altX, sameX, interaction1 = "gender_RS") # model1
# include effect of h index later too etc
# look up which effects to include




myeff <- includeEffects(myeff, egoX, altX, inPopX, interaction1 = "h_index_RS") # model 2 (as long as myeff is eithin the brackets too, the other things get added to the myeff object)


# change statistics to fit whatever I want to know about h-index 
# not sameX 
# look up inPopX: preferential attachment (cf. Koen lel), higher h-index more popular 
  # cf. Fridas modeling process (Jochem told her that )

#diffXTransTrip
#different covariate × transitive triplets (diffXTransTrip), defined by the number of transitive triplets i → h → j ← i that have different covariate values for i and j, diff. cov. values = diff gender -RSiena manual 

# "The transtrip for different gender (diffXTransTrip) is positive and significant, which implies that an ego prefers to close triads with alters who have a different gender score than theirs. This is an interesting finding, because it implies that their is a tendency towards diverse collaboration. However, when compared to the level of gender clustering and the fact that this clustering decreased, this finding does make sense." - Niels


```

```{=html}
<!-- 


hint FOR MY OWN PROJECT: interaction effect of these statistics and the
effect of gender

THE "X" in altX and eogX refer to the "interaction 1" variable, so it's eg. ego x gender_RS and alter x gender_RS!

altX - egos send more ties who have more on gender (aka. gender =1) 
(iiiish look at egoX explanantion, this is basically the same but in "reverse", aka about receiving rather than sending)

-   covariate-alter or covariate-related popularity (altX), defined by
    the sum of the covariate over all actors to whom ***i*** has a tie,
-   same as egoX but RECEIVING ties

egoX - [definition?]

-   covariate-ego or covariate-related activity (egoX), defined by i’s
    out-degree weighted by his covariate value,
-   since male = 1 egoX only "looks at" men's ties 
-   if egoX significant then men's ties sign. diff. based on whether its to a man or woman (look at direction of effect to know which gender) 



sameX (Gender): Tests whether scholars prefer same-gender collaborators (direct measure of gender homophily).
model the general tendency for scholars to collaborate with same-gender peers.

  Confirmation:

    Significant and positive coefficient for sameX in the RSiena output would confirm that gender homophily exists.

    A non-significant or negative coefficient would disconfirm this hypothesis.


    

"interaction1" HAS NOTHING TO DO WITH AN INTERACTION EFFECT

-   If there are several dependent variables (networks and/or behavioral
    variables), the name of the dependent variable (name) also is
    required to specify the effect, and will be given in the column
    interaction1 (etc.) of the effects object.

-   It should be mentioned that the keywords interaction1, interaction2,
    and interaction3 in functions such as includeEffects do not point to
    interactions between effects. These keywords are used only to
    exactly specify an effect. They point to the covariate or other
    dependent variable used in the effect; this name was chosen for the
    keyword because this other variable may be said to be interacting
    with the dependent variable (network or behavior).

-->
```
<!-- ### test 4: Specify model/algorithm -->

```{r algo, hide =TRUE, echo=FALSE, eval=FALSE}
# Algorithm from when Jochem and I sat together but it's not fitting anymore bc model type is for undirected ties 
# myAlgorithm <- sienaAlgorithmCreate(modelType = c(nets=6), projname = "proProject")
# undirected: model 6 "additive model", #check manual for this NOW IT'S UNDIRECTED!!! 

# Niels advice: take his algorithm instead. haha ok, no questions asked my guy, you got it. 
myAlgorithm_test <- sienaAlgorithmCreate (projname = "results_test")

```

<!-- ### test 5: Estimate model -->

```{r estimate, hide =TRUE, echo=FALSE, eval=FALSE}
# estimate the model 
model_test <- siena07(myAlgorithm_test, data = mydata, effects = myeff, returnDeps = TRUE)
#latter needed for GOF lateron

model_test

```

```{=html}
<!-- 
output:  convergence ration NA probably bc we chose all ties to be included (type = "all") earlier
we can go back and change that or change it here in the myAlgorithm function

convergence ratio should be <.25   !!!
- if above, run model again
- if stays above .25, remove some of the netw. stats
- if close to threshold (e.g. 0.2584) add prevAns = model1  to the siena07 command and rerun then
      model1 <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns= model1, returnDeps = TRUE)

-->
```
<!-- ### test 6: GOF - Goodness of Fit -->

```{r GOF test, hide =TRUE, echo=FALSE, eval=FALSE}
GOF <- sienaGOF(model_test, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF, main = "Test Indegree Distr")


# "Error using packet 1 "bw" is not positive"

#--------- Try to fix error:  -------  
##--------- Inspect GOF object  ------- 
# str(GOF)
  # summary(GOF$Joint$SimulatedTestStat)

##--------- Check for invalid /infinte values ------- 
    # anyNA(GOF$Joint$SimulatedTestStat)  # Should return FALSE
    # any(!is.finite(GOF$Joint$SimulatedTestStat))  # Should return FALSE
    # summary(GOF$Joint$SimulatedTestStat)  # Inspect the range of values


    # Get rid of invalid or infinite values
        # GOF$Joint$SimulatedTestStat <- na.omit(GOF$Joint$SimulatedTestStat)
        # GOF$Joint$SimulatedTestStat <-   GOF$Joint$SimulatedTestStat[is.finite(GOF$Joint$SimulatedTestStat)]

  # var(GOF$Joint$SimulatedTestStat)  # Should return a positive value


##--------- Density simulation  ------- 
    # simulated <- GOF$Joint$SimulatedTestStat
    # # Try density estimation with default bandwidth
    #   density_est <- tryCatch(density(simulated), error = function(e) e)
    #   if (inherits(density_est, "error")) {
    #   print("Density estimation failed.")
    #     } else {
    #             print("Density estimation succeeded.")
    #             plot(density_est, main = "Simulated Test Statistics")
    #     }
    #   
##--------- Force bandwidth -------     
      # Force the Bandwidth in plot.sienaGOF
      # plot(GOF, main = "Indegree Distribution Goodness of Fit", bw = 0.5)
      # #nope does not work 

##--------- Custom function -------  
      # custom_plot_sienaGOF <- function(gof, bw = 0.5, ...) {
      # simulated <- gof$Joint$SimulatedTestStat
      # observed <- gof$Joint$ObservedTestStat
      # 
      # # Perform density estimation with forced bandwidth
      # density_est <- density(simulated, bw = bw)
      # 
      # # Plot the density
      # plot(density_est, main = "Custom Indegree Distr.- GOF", ...)
      # abline(v = observed, col = "red", lty = 2, lwd = 2) 
      # }
      # 
      #  #run custom function      
      # custom_plot_sienaGOF(GOF, bw = 0.5)
      # # works but not the same at all 
      # 
##--------- Ggplot2 ------- 
# Fuck this shit. I using ggplot 
    # library(ggplot2)
    # 
    # # Prepare data
    # data <- data.frame(
    #   value = GOF$Joint$SimulatedTestStat,
    #   type = "Simulated"
    # )
    # data <- rbind(data, data.frame(value = GOF$Joint$ObservedTestStat, type = "Observed"))
    # 
    # # Plot density
    # ggplot(data, aes(x = value, fill = type)) +
    #   geom_density(alpha = 0.5, adjust = 1, bw = 0.5) +
    #   geom_vline(xintercept = GOF$Joint$ObservedTestStat, color = "red", linetype = "dashed") +
    #   ggtitle("Indegree Distribution Goodness of Fit") +
    #   theme_minimal()

# works but it's the same thing as with the custom function... It not a violin-plot and just does not hit the spot. 
# The only violins I have are the ones playing the sad self-pity music in my brain right now. 

##--------- reinstall RSiena ------- 

# I.... 
    # ....
      # .... 
        # .... sure.  


# --------- I hate myself and back we go to GOF -----
# other siena GOF functions (outdegree does not work either, GeodesicDistribution weirdly enough works.)

#gofo0 <- sienaGOF(model_test, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20), varName = "nets")
#plot(gofo0, main ="Test Outdegree Distr")

#gof0.gd <- sienaGOF(model_test, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "nets")
#plot(gof0.gd, main ="Test Geodesic")


```

```{=html}
<!-- 

Make it a directed network:  ties from "all" to "first"
focus: writing,
model interpretation


### Questions

-   Is the use of RSiena different for people who look at 1 vs 2 waves.
    I need to look at one, made 2 anyways bc RSiena needs waves to do
    it's thing but:\
    Doesn't it look at ministeps over time? How can I interpret the
    output in a way that pertains to only one wave when the process
    seems to consider it within the construct of 'future change'?

    -   How does my model interpretation potentially differ from peeps
        who looked at two waves (or more)?

-   not looking at TWoStep! diff. modelling process. 
RSiena works with ministeps indeed not just Twosteps (that works with two steps lol)


- diff. betweeen wave 1 and wave 2 model -> convergence 
- default recip. and degree: estimates the putput based on a certain reciproc. and degree "assumption"
 
-->
```
## \<\> Model 0 - Null Model

First, I estimated a null model with only reciprocity and degree. The
maximum convergence ratio of the model is 0.084. The GOF on the Indegree
Distribution is zero**.** This indicates that the model is a very bad
fit for the data and estimates of this model do not give us reliable
information. \
*For all GOF plots compiled in one place, see here:
<https://treesdude.github.io/labjournal/finalGOF.html>.\
*For the sake of practise, here they are anyways:\
The outdegree (density) effect assesses the tendency of nodes to
establish outgoing connections, serving as a baseline for the likelihood
that any given node will initiate an outgoing tie. In the context of
this study, these outgoing ties represent the initiation of
co-publications. The outdegree estimate is significant and negative (b =
-2.6388, SE = 0.1460), suggesting that a scholar's tendency to initiate
collaboration is low and dense networks are less sought after. The
reciprocity parameter id significant and positive (b = 1.8035, SE =
0.4343), which implies that directed ties are likely to be reciprocated.
Specific to our case, it describes that scholars tend to initiate
collaborations with those alters who sent a tie to them too.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m0}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M0 <- sienaDataCreate(nets)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M0, modelname = "./results/results_M0") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff M0}
# define effects structure 
myeff_M0 <- getEffects(mydata_M0)

```

### ---Step 3: Add Effects

Null model: none.

### ---Step 4: Specify model/algorithm

```{r algo M0}

myAlgorithm_M0 <- sienaAlgorithmCreate (projname = "results_M0")

```

### ---Step 5: Estimate model

```{r estimate M0, eval=FALSE}
# estimate the model 
Model0 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, returnDeps = TRUE)
Model0

fsave(Model0, file ="estimated_M0")
```

```{r load M0 output, echo=FALSE}
Model0 <- fload("./data/processed/estimated_M0_20241123.rda")
Model0
```

```{r GOF M0}
GOF_M0 <- sienaGOF(Model0, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M0, main = 'GOF - "Null" Model 0 \n Indegree Distribution')
```

------------------------------------------------------------------------

## \<\> Model 1 - Structural Effects

This model additionally entails InPop, outAct, isolateNet to control for
effects of network statistics. The maximum convergence ratio of this
model is **0.17** and a Goodness of Fit on the indegree distribution is
**0.002.** Estimates of this model do not fit the data well and thus do
not give us reliable information.

For the sake of practise, here they are anyways:\
As in our null model, the parameters for outdegree (b = -2.0862, SE =
0.7050) and reciprocity (b = 1.0894, SE = 0.4144) remain significant and
unchanged on their 'direction'.

The added parameter for indegree popularity tells us something about the
popularity of people who have a lot of outgoing ties. In our model the
estimate for this parameter is significantly positive (b =0.2549, SE
=0.0439), suggesting that scholars specifically like to initiate
collaborations with those who have a lot of indegrees.

Further, the parameter for outdegree activity conveys information about
scholars' preferences to collaborate with others who have a lot of
outdegrees. In our case, the estimate is insignificant and cannot be
reliably interpreted.

Lastly, the network isolate parameter can be used to infer about
scholars' tendency to neither send nor receive ties. The estimate for
this parameter is significantly positive (b = 4.9645, SE = 1.1146) in
our model, indicating that the scholars of our sample tend to remain
isolates.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define M1}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M1 <- sienaDataCreate(nets)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M1, modelname = "./results/results_M1") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m1}
# define effects structure 
myeff_M1 <- getEffects(mydata_M1)

```

### ---Step 3: Add Effects

```{r effects m1}

myeff_M1 <- includeEffects(myeff_M1, isolateNet, inPop, outAct)

```

### ---Step 4: Specify model/algorithm

```{r algo m1}

myAlgorithm_M1 <- sienaAlgorithmCreate (projname = "results_M1")

```

### ---Step 5: Estimate model

```{r estimate m1, eval = FALSE}
# estimate the model 
Model1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, returnDeps = TRUE)
Model1
# maximum convergence >.25
# run again:
Model1.1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, prevAns= Model1, returnDeps = TRUE)
Model1.1

#fsave(Model1, file ="estimated_M1")
#fsave(Model1.1, file = "estimated_M1.1")

```

```{r load M1 output, echo=FALSE}

Model1 <- fload("./data/processed/estimated_M1_20241123.rda")
Model1

Model1.1 <- fload("./data/processed/estimated_M1.1_20241123.rda")
Model1.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M1}
GOF_M1.1 <- sienaGOF(Model1.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M1.1, main = "GOF - Model 1 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 2 - Gender

This model is partly similar to Model 1 but additionally holds the
network statistics sameX, diffXTransTrip, egoX & altX of the covariate
gender. EgoX and AltX ought to indicate whether male or female scholars
are more likely to initiate or receive ties, while sameX models the
general tendency for scholars to collaborate with same-gender peers.

The statistic for transitive triplets based on a different covariate
value (here: different gender) was originally included to control for
relationships of triadic nature (diffXTransTrip), but will be
disregarded in the interpretation of the results since a more
appropriate control statistic would have been that of transitive
triplets (TransTrip), but I simply lack the time to adjust the models.

```         
Estimates, standard errors and convergence t-ratios                                                
                                            Estimate   Standard   Convergence
                                                        Error      t-ratio     
Rate parameters:    
0       Rate parameter                       2.9204  ( 0.4617   )               

Other parameters:    
1. eval outdegree (density)                 -1.8628  ( 1.1060   )    0.0703 --> insign. (-1.684268)    
2. eval reciprocity                          1.3341  ( 0.4543   )   -0.0152 --> sign.    (2.936606)***     
3. eval indegree - popularity                0.2209  ( 0.0470   )   -0.0083 --> sign.    (4.7)     ***
4. eval outdegree - activity                -0.3017  ( 0.2287   )    0.0806 --> insign. (-1.319195)    
5. eval network-isolate                      4.7507  ( 1.5194   )   -0.0379 --> sign.    (3.126695)***     

6. eval gender_RS alter                      0.3254  ( 0.2701   )    0.0411 --> insign.  (1.204739)      
7. eval gender_RS ego                       -1.0642  ( 0.6055   )   -0.0400 --> insign. (-1.757556)     
8. eval same gender_RS                       0.6637  ( 0.2685   )    0.0308 --> sign.    (2.471881)***     
9. eval transitive triplets diff. gender_RS  2.1787  ( 0.6027   )    0.0416 --> sign.    (3.6149)  ***
```

The maximum convergence ratio for model 2 is **0.132** and the Goodness
of Fit on the indegree distribution is **0.096**.

Looking at our 'base-line' network statistics, it bears noting that the
estimate for Outdegree is not significant anymore. This statistic's
prior significance can seemingly be explained by a parameter relating to
the covariate gender that I added to this model. However, the estimate
of the outdegree parameter is still nearing the significance 'threshold'
(b = -1.8628, SE = 1.1060, b/SE = -1.684268).

Comparing this model to the prior ones, we see that the reciprocity
estimate is still significantly positive (b = 1.3341, SE = 0.4543).

The parameters of the structural network statistics also show similar
patterns in estimates and significance as they did in model 1.

Concerning the added covariate "gender", we find that the parameter
estimates for both covariate-related activity (egoX) , pertaining to the
gender of the ego, and covariate-related popularity (altX), relating to
the gender of the alter, do **not** show any significance and hence do
not allow for reliable inferences.

The 'same-covariate' measure (sameX) for gender is significantly
positive (b = 0.6637, SE =0.2685). It reveals that scholars, indeed,
prefer same-gender collaborators. This supports H1, which posits gender
homophily in collaboration patterns.

Model 2 provides strong evidence for gender homophily (H1), with
scholars more likely to collaborate with same-gender peers, but no
significant differences in collaboration tendencies based on the gender
of ego or alter.

Hypothesis about structural effects

-   **Network Statistics**:

    -   **sameX (Gender)**: Tests for gender homophily directly.

    -   Structural statistics like **reciprocity**, **inPop**,
        **outAct**, and **isolateNet** are control variables to account
        for network-level dynamics.

-   **Confirmation**:

    -   A significant `sameX` coefficient, even after including
        structural controls, would confirm that gender homophily exists
        at the dyadic level.

    -   If `sameX` becomes non-significant after controls, it suggests
        structural factors, not individual preferences, explain gender
        clustering.

--\> unsure if this hpold true for me, since I added gender only after
running a model that controls for structural effects so I cannot really
infer that but it'S too late to change that now.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m2}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

# define data object 
mydata_M2 <- sienaDataCreate(nets, gender_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M2, modelname = "./results/results_M2") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m2}
# define effects structure 
myeff_M2 <- getEffects(mydata_M2)
```

### ---Step 3: Add Effects

```{r effects m2}
myeff_M2 <- includeEffects(myeff_M2, isolateNet, inPop, outAct) 

myeff_M2 <- includeEffects(myeff_M2, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") 
```

```{=html}
<!--
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collaborate with people who are popular

# sameX
# same covariate, which can also be called covariate-related identity (sameX), defined by the number of ties of i to all other actors j who have exactly the same value on the covariate, s net i95(x) = P jxij I{vi = vj},
# where the indicator function I{vi = vj} is 1 if the condition {vi = vj} is satisfied, and 0 if it is not;


#diffXTransTrip
different covariate × transitive triplets (diffXTransTrip), defined by the number
of transitive triplets i → h → j ← i that have different covariate values for i and j,
diff. cov. values = diff gender

-->
```
### ---Step 4: Specify model/algorithm

```{r algo m2}

myAlgorithm_M2 <- sienaAlgorithmCreate (projname = "results_M2")

```

### ---Step 5: Estimate model

```{r estimate m2, eval =FALSE}
# estimate the model 
Model2 <- siena07(myAlgorithm_M2, data = mydata_M2, effects = myeff_M2, returnDeps = TRUE)
Model2
# maximum convergence >.25
# run again:
Model2.1 <- siena07(myAlgorithm_M2, data = mydata_M2, effects = myeff_M2, prevAns= Model2, returnDeps = TRUE)
Model2.1 

#fsave(Model2, file ="estimated_M2")
#fsave(Model2.1, file = "estimated_M2.1")

```

```{r load M2 output, echo=FALSE}

Model2 <- fload("./data/processed/estimated_M2_20241123.rda")
Model2

Model2.1 <- fload ("./data/processed/estimated_M2.1_20241123.rda")
Model2.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M2}
GOF_M2.1 <- sienaGOF(Model2.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M2.1, main = "GOF - Model 2 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 3 - Department

This model is partly similar to Model 2 but additionally holds the
network statistics egoX and altX for the main effect of department, to
investigate differences in same-gender collaboration patterns based on
the two departments.

Originally, I planned to also include an interaction term of departm_RS
and gender_RS to obtain clearer insights into variations in patterns of
gender homophily in relation to scholars' departments. However, I have
refrained from adding another model to this project. I decided against
this partly due to time constraints but I also reason that I can use the
comparison of model 2 and 3 to examine changes in the estimate of
'sameX' (gender), to *tentatively* infer whether the department of
scholars modifies the collaboration patterns at hand.

Model 3 has a maximum convergence ratio of **0.182** and a Goodness of
Fit on the indegree distribution of **0.108**.

Interestingly, the estimate for outdegree (density) that had lost its
significance in model 2, is now back to being significant and negative
(b = -4.3233, SE = 1.0871).

In terms of estimate size and significance of the structural network
statistics we, once again, find similar patterns as in models 1 and 2.
However, the outdegree parameter has now reached significance (b =
-4.3233, SE = 1.0871).

With regard to the network statistics of the covariate 'gender', we find
that the estimates relating to the gender of an ego (egoX) and the
gender of an alter (altX) remain non-significant. The parameter
estimates for sameX remains significantly positive (b = 0.6632, SE =
0.2623) and are almost identical to those of model 2 (b = 0.6632, SE =
0.2623). This suggests that adding the covariate 'department' did not
bring about any noteworthy changes in patterns of scholars seeking out
same-gendered scholars to collaborate.

When it comes to the parameters pertaining to the added covariate
'department', we find that any differences in collaboration based on the
department of an ego (egoX) or of an alter (altX) are not warranted, as
the estimates of these parameters are insignificant. However, we find a
significantly positive estimate for the parameter of 'covariate-related
identity' (sameX), which indicates that scholars are more likely to
collaborate with people from their own department, indicating
**department-based homophily**. (b = 2.7240, SE = 0.7277).

Model 3 highlights persistent gender homophily (H1) across departments
but does not provide evidence for differences in gender homophily
between Sociology and Political Science, contrary to H2.

------------------------------------------------------

Estimates and SEs model 3:

1\. eval: outdegree (density) -4.3233 ( 1.0871) \*\*\*

2\. eval: reciprocity 1.0725 ( 0.4174) \*\*\*

3\. eval: indegree - popularity 0.2122 ( 0.0517) \*\*\*

4\. eval: outdegree - activity -0.3364 ( 0.2182) insign. (-1.541705)

5\. eval: network-isolate 4.2006 ( 1.3007) \*\*\*

6\. eval: gender_RS alter 0.3356 ( 0.2659) insign (1.262129)

7\. eval: gender_RS ego -1.0066 ( 0.5121) insign (-1.965632) almost
significant

8\. eval: same gender_RS 0.6632 ( 0.2623) sign. (2.52840) \*\*\* \
(sameX: comparison model 2 and 3:\
estimate got hardly any smaller (0.6637--\>0.6632), SE got hardly any
smaller (0.2685--\>0.2623), significance value got bigger
(2.47881--\>2.52840))

9\. eval: tr. tripl. diff. gend. 1.5479 ( 0.5403) sign. (2.86489)
\*\*\*\
(diffxTrTrip comparison model 2 and 3:\
estimate got smaller and NEGATIVE (2.1787 –\>-1.5479), SE got smaller
(0.6027 –\> 0.5403), significance value got smaller (3.6149--\>2.86489))

10\. eval: departm_RS alter 0.4141 ( 0.7546) insign. (0.549)

11\. eval: departm_RS ego -1.1573 ( 0.9148) insign (-1.27)

12\. eval: same departm_RS 2.7240 ( 0.7277) \*\*\*

-   **Network Statistics**:

    -   **sameX (Gender)**: Compute and compare the size and
        significance of the coefficient for each department. This
        requires separate models for each department or interaction
        terms that allow the effect to vary by department.

-   **Confirmation**:

    -   A significantly stronger positive coefficient for `sameX` in
        Political Science compared to Sociology would confirm the
        hypothesis.

    -   Similar coefficients or reversed patterns would disconfirm the
        hypothesis.

altX - egos send more ties who have more on gender (aka. gender =1)
(iiiish look at egoX explanantion, this is basically the same but in
"reverse", aka about receiving rather than sending)

-   covariate-alter or covariate-related popularity (altX), defined by
    the sum of the covariate over all actors to whom ***i*** has a tie,
-   same as egoX but RECEIVING ties

egoX - [definition?]

-   covariate-ego or covariate-related activity (egoX), defined by i’s
    out-degree weighted by his covariate value,
-   since male = 1 egoX only "looks at" men's ties
-   if egoX significant then men's ties sign. diff. based on whether its
    to a man or woman (look at direction of effect to know which gender)

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m3}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

#put department in object 
department <- factor(df_ego$discipline.24, levels = c("sociology", "political science"))
#make numeric
department_numeric <- as.numeric(department) - 1 # sociology = 0, political science = 1
# covariate 2
departm_RS <- coCovar(department_numeric)

# define data object 
mydata_M3 <- sienaDataCreate(nets, gender_RS, departm_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M3, modelname = "./results/results_M3") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m3}
# define effects structure 
myeff_M3 <- getEffects(mydata_M3)
```

### ---Step 3: Add Effects

```{r effects m3}
myeff_M3 <- includeEffects(myeff_M3, isolateNet, inPop, outAct) # model 1

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") #model 2

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, sameX, interaction1 = "departm_RS") #model 3

```

```{=html}
<!--
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collaborate with people who are popular

# sameX
# same covariate, which can also be called covariate-related identity (sameX), defined by the number of ties of i to all other actors j who have exactly the same value on the covariate, s net i95(x) = P jxij I{vi = vj},
# where the indicator function I{vi = vj} is 1 if the condition {vi = vj} is satisfied, and 0 if it is not;


#diffXTransTrip
different covariate × transitive triplets (diffXTransTrip), defined by the number
of transitive triplets i → h → j ← i that have different covariate values for i and j,
diff. cov. values = diff gender

-->
```
### ---Step 4: Specify model/algorithm

```{r algo m3}

myAlgorithm_M3 <- sienaAlgorithmCreate (projname = "results_M3")

```

### ---Step 5: Estimate model

```{r estimate m3, eval =FALSE}
# estimate the model 
Model3 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, returnDeps = TRUE)
Model3

#if maximum convergence >.25: run again
Model3.1 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, prevAns= Model3, returnDeps = TRUE)
Model3.1 

#fsave(Model3, file ="estimated_M3")
#fsave(Model3.1, file = "estimated_M3.1")

```

```{r load M3 output, echo=FALSE}

Model3 <- fload("./data/processed/estimated_M3_20241123.rda")
Model3

Model3.1 <- fload ("./data/processed/estimated_M3.1_20241123.rda")
Model3.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M3}
GOF_M3.1 <- sienaGOF(Model3.1 , IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M3.1, main = "GOF - Model 3 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 4 - H-Index

This model further builds on the prior model and also includes network
statistics of the covariate H-Index, namely egoX, altX and inPopX. By
adding these effects we can model how the h-index of a scholar
influences their preferences regarding same-gendered collaborations.
Similarly, to model 3, this model could have profited from a 'follow-up'
model that includes and interaction term between h-index and gender to
better infer about their influences. The same reasoning for my decision
to stick with one model prior also applies to this case. EgoX for this
covariate examines how the h-index of a scholar affects their
collaboration behavior, while AltX examines how the h-index of potential
collaborators affects their attractiveness as partners. With the help of
these effects we can test prestigious scholars are more likely to
initiate or receive ties. By looking at in-degree popularity weighted by
sender’s H-index (inPopXi) we can assesse whether a high h-index
increases the likelihood of being sought out as a collaborator.\

```         
Estimates, standard errors and convergence t-ratios
                                               Estimate   Standard   Convergence
                                                          Error      t-ratio
Rate parameters:  
  0      Rate parameter                       2.6891  ( 0.3675   )               

Other parameters:     
   1. eval outdegree (density)                 -2.7693  ( 1.9205   )    0.0698 --> insign. (-1.441968)       
   2. eval reciprocity                          2.1770  ( 0.5960   )    0.0919 --> sign.    (3.652685)***      
   3. eval indegree - popularity               -0.0653  ( 0.1654   )   -0.0104 --> insign. (-0.3948005)      
   4. eval outdegree - activity                -0.6851  ( 0.4029   )    0.0527 --> insign. (-1.700422)      
   5. eval network-isolate                      5.4634  ( 2.2770   )    0.0028 --> sign.    (2.399385)***      
   6. eval gender_RS alter                     -0.0903  ( 0.3372   )    0.0129 --> insign. (-0.2677936)       
   7. eval gender_RS ego                        0.2993  ( 0.4934   )   -0.0503 --> insign. (0.6066072)      
   8. eval same gender_RS                       0.6820  ( 0.2804   )    0.0201 --> sign.   (2.43224)  ***       
   9. eval transitive triplets diff. gender_RS  2.4871  ( 0.6670   )    0.0505 --> sign.   (3.728786) ***    
  10. eval departm_RS alter                     0.3929  ( 0.6128   )   -0.0061 --> insign. (0.6411554)    
  11. eval departm_RS ego                      -1.4652  ( 0.6963   )   -0.0115 --> SIGN.   (-2.104265)***
  12. eval same departm_RS                      2.6327  ( 0.6080   )    0.0173 --> sign    (4.330099) ***    
  13. eval h_index_RS alter                     0.0829  ( 0.0364   )   -0.0031 --> SIGN    (2.277473) ***
  14. eval h_index_RS ego                      -0.1591  ( 0.0744   )    0.0161 --> SIGN.  (-2.138441) ***   
  15. eval ind. pop.^(1/1) weighted h_index_RS -0.0287  ( 0.0570   )   -0.0129 --> insign.(-0.5035088)
```

Model 4 has a maximum convergence ratio of **0.23,** and a a Goodness of
Fit on the indegree distribution of **0.197**.

Structural network statistics:

The estimate for outdegree is negative but not significant. This
suggests that, after accounting for other variables, the baseline
likelihood of forming ties remains low and is not a major driver of
collaboration patterns. Reciprocity has a significant positive estimate
(b = 2.1770, SE = 0.5960). This reflects a strong tendency for mutual
collaborations, meaning if one scholar collaborates with another, the
latter is likely to collaborate back (dyadic reciprocity). The paramters
Indegree Popularity and Outdegree Activity are both insignificant,
suggesting that neither the tendency to receive many collaborations
(indegree popularity) nor the tendency to initiate them (outdegree
activity) is significantly relevant to the network structure. The
network isolate estimate is, once again, significantly positive ( b =
5.4634, SE = 2.2770), indicating that as isolates tend to remain
isolated.

Gender:

Similar to AltX and EgoX in prior models, both an alter's gender (b =
-0.0903, SE = 0.3372) and an ego's gender (b = 0.2993, SE = 0.4934)
still have insignificant effects. The parameter SameX shows a
significant positive estimate (b = 0.6820, SE = 0.2804) which barely
differs from the two prior models. This suggests that adding the
covariate 'H-Index' did not prompt any major changes in scholars'
preferences for same-gendered collaborators.

Department:

While an alter's departmen**t** (AltX) remained insignificant to the
collaboration patterns, the estimates for an ego's department changed
incomparison to the prior model. In this model, egoX seems to be of
significant relevance (b = -1.4652, SE = 0.6963), with its negative
estimate indicating that, once effects of h-index are controlled for,
scholars from the political science (coded as 1) department have a lower
tendency to initiate collaborations than scholars from the sociology
department (coded as 0). \
Similar to the prior model, the sameX parameter for department shows a
significant positive estimate (b = 2.6327, SE = 0.6080) scholars are
still more likely to collaborate with others from the same department.

H-Index:

Both an alter's and an ego's H-Index seem to be relevant in
collaboration preferences. The significant and positive estimateof AltX
(b = 0.0829, SE = 0.0364) suggest that scholars with higher h-index are
more preferred as collaborators. The egoX estimate is significant and
negative (b = -0.1591, SE = 0.0744) suggesting that scholars with higher
h-index are less likely to initiate collaborations.

*Discussion part about egoX h.index: **"This is in line with theories of
preferential attachment and upward mobility, because as egos with higher
prestige will are more likely to receive request for collaboration
rather then sending request for collaboration." - Niels***

The estimate for the Indegree Popularity weighted by h-index is
insignificant, suggesting that high-prestige individuals (as measured by
their h-index) do not significantly receive more indegrees

The results partially support **H4**, as the h-index is shown to
influence collaboration patterns, with prestigious scholars being more
likely to attract collaborations as alters but less likely to initiate
them as egos. However, gender homophily remains significant even after
controlling for the h-index, suggesting that prestige does not
substantially weaken the influence of gender homophily in co-authorship
networks.

Transitive Triplets (Diff Gender): A positive significant effect
(b=2.4871, SE = 0.6670). This suggests that transitive structures (where
two people connected to a common third party collaborate) are more
likely even across gender differences. Could be a potential
counterbalance to homophily.

R-Manual: in-degree popularity weighted by sender’s V (inPopX)
defined by the sum of values of V for actors h for which there are ties
i → j ← h

note that to use this effect for p = 2 the variable V must be
nonnegative, which
implies that it must be non-centered;
further note that if V is a behavioral variable, it will be used in the
centered version,
so that p must be 1; if weighting by a non-centered behavioral variable
is desired, a
new effect is necessary;

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m4}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender
# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

#put department in object 
department <- factor(df_ego$discipline.24, levels = c("sociology", "political science"))
#make numeric
department_numeric <- as.numeric(department) - 1 # sociology = 0, political science = 1
# covariate 2
departm_RS <- coCovar(department_numeric)

# put h-index in object 
h_index <- df_ego$h_index
df_ego$h_index
# covariate 3
h_index_RS <- coCovar(h_index) # h_index as covariate in RSiena

# define data object 
mydata_M4 <- sienaDataCreate(nets, gender_RS, departm_RS, h_index_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M4, modelname = "./results/results_M4") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff M4}

# define effects structure 
myeff_M4 <- getEffects(mydata_M4)

```

### ---Step 3: Add Effects

```{r effects M4}

myeff_M4 <- includeEffects(myeff_M4, isolateNet, inPop, outAct) # model 1

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") #model 2

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, sameX, interaction1 = "departm_RS") #model 3
  
myeff_M4 <- includeEffects(myeff_M4, egoX, altX, inPopX, interaction1 = "h_index_RS") # model 4

```

```{=html}
<!--
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collab with people who are popular

# sameX
# same covariate, which can also be called covariate-related identity (sameX), defined by the number of ties of i to all other actors j who have exactly the same value on the covariate, s net i95(x) = P jxij I{vi = vj},
# where the indicator function I{vi = vj} is 1 if the condition {vi = vj} is satisfied, and 0 if it is not;

# look up inPopX: preferential attachment (cf. Koen lel), higher h-index more popular 
  # cf. Fridas modeling process (Jochem told her that )

-->
```
### ---Step 4: Specify model/algorithm

```{r algo M4}

myAlgorithm_M4 <- sienaAlgorithmCreate (projname = "results_M4")

```

### ---Step 5: Estimate model

```{r estimate M4, eval =FALSE}
# estimate the model 
Model4 <- siena07(myAlgorithm_M4, data = mydata_M4, effects = myeff_M4, returnDeps = TRUE)
Model4

#fsave(Model4, file ="estimated_M4")

```

```{r load M4 output, echo=FALSE}

Model4 <- fload("./data/processed/estimated_M4_20241123.rda")
Model4

```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M4}
GOF_M4 <- sienaGOF(Model4, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M4, main = "GOF - Model 4 \n Indegree Distribution")
```
