---
title: "RQdraft2"
author: "Verooo"
date: "2024-09-18"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

------------------------------------------------------------------------

# Research Questions 

RQ1a: To what extent does gender homophily occur across all co-publications of the staff of the Sociology Department and Political Science Department at Radboud University?

RQ1b: To what extent do the two departments differ in their gender homophily?

RQ2: How to does gender homophily affect collaboration differently for collaborations between staff who have published in low vs. high prestige journals?

RQ3: To what extent do structural network effects influence the gender homophily of each department?

## Horta et al. (2022)

-   male researchers preferred to co-publish with other male researchers,

-   and female researchers preferred to collaborate with other females.

-   researchers of approximately similar ages tended to collaborate more with each other.

-   geographical, cultural, & institutional proximity were strongest predictors of collaboration.

-   analysis shows that in the social sciences, ACQUIRED attributes (particularly preferences regarding research agendas) take precedence over ASCRIBED attributes (e.g. gender or age)

-   Horta et al. (2022) - RQs:

    -   

        (1) What homophily preferences determine research collaborations among higher education researchers?

    -   

        (2) What are the most powerful homophily indicators in research collaborations in higher education research?

## Kwiek & Roszka (2021)

-   Most male scientists collaborate solely with males.
-   Most female scientists, in contrast, do not collaborate with females at all.
-   All-female authorship marginal, all-male authorship pervasive, in all age groups.
-   Articles written in mixed-sex collaboration published in more prestigious journals.

------------------------------------------------------------------------

# References:

Horta, H., Feng, S., & Santos, J. M. (2022). Homophily in higher education research: a perspective based on co-authorships. Scientometrics, 127(1), 523–543. <https://doi.org/10.1007/s11192-021-04227-z>

Kwiek, M., & Roszka, W. (2021). Gender-based homophily in research: A large-scale study of man-woman collaboration. Journal of Informetrics, 15(3), 101171. <https://doi.org/10.1016/j.joi.2021.101171>

# Data types needed for research

+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Variable                          | Potential measurement                                   | Notes                                                                                                  |
+===================================+=========================================================+========================================================================================================+
| **Actor attributes**              |                                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Name                              | Name from university department website                 | using "read_html()" function                                                                           |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Age                               | length of career                                        | time since first publication                                                                           |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Gender                            | Google Scholar:\                                        | What about gender neutral names? If we go by this principle, then gender neutral names = non-binary ?? |
|                                   | Name ➙ Look up in gendered list whether Woman or Man    |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
|                                   | Linked-in:\                                             | Not everyone puts their pronouns on linkedin\                                                          |
|                                   | Pronouns                                                | And: not everyone who uses e.g. She/her pronouns is a Woman                                            |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| High prestige publication (prior) | If the author had a publication in a journal that ....\ |                                                                                                        |
|                                   | before the publication of the paper in question         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Affiliation                       | Name of university they work at (radboud university)    | Taken from university website about department                                                         |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| **Relational Attributes**         |                                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Initiation of the co-authorship   | First author of paper = initiator                       |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Paper                             | A paper people have co-authorship on                    | How prevent from counting double?\                                                                     |
|                                   |                                                         | Overall: systematic approach to webscrape?                                                             |
|                                   | Signifier: DOI?                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
|                                   |                                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+

: Data types
